---
title: "2023 MLB Batters Data Analysis Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: sandstone
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

<style type="text/css">
/* Get a fancy font from Google Fonts */
@import url('https://fonts.googleapis.com/css2?family=Yusei+Magic&display=swap');

body {
  background-color: lightgray;
  color: darkgreen; /* text color */
  text-align: left;
}

.navbar {
	position: fixed;
	display: flex;
	justify-content: center;
	width: 100%;
}

/* Make text visible on inputs */
.shiny-input-container {
  color: purple;
}
<!--
.navbg {
  background-color: blue;
  color: white;
}
-->
</style>                    

\

```{r setup, include=FALSE}
if (!require("flexdashboard")) {
   install.packages("flexdashboard")
   library(flexdashboard)
}
if (!require("shiny")) {
   install.packages("shiny")
   library(shiny)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("stats")) {
   install.packages("stats")
   library(stats)
}
if (!require("fANCOVA")) {
   install.packages("fANCOVA")
   library(fANCOVA)
}
```


```{r data}
batters = read.csv(file="https://raw.githubusercontent.com/azheneghan/aheneghan/main/datasets/mlb_Batters.csv")[,-1]
batters <- batters %>%
  dplyr::rename(RunsScored = R, NumberofHits = H, NumberofHomeRuns = HR, NumberofWalks = BB, NumberofStrikeOuts = SO, AtBats = AB)
y.names = names(batters)[c(5,6,7,10,14,15)]
x.names = names(batters)[c(5,6,7,10,14,15)]
team = unique(batters$Team)
```


\


Column {.sidebar data-width=200}
-------------------------------------------------------------------

\

<hr>

<!-- put all input widgets in the same R code chunk.
     The code is the same as we used in the regular R shiny apps-->
```{r}
selectInput(inputId="Team", 
            label = strong("Team"), 
            choices = c("All", as.vector(team)), 
            selected = "All")

selectInput(inputId="Y", 
            label = strong("Response Variable: Y"), 
            choices = y.names[], 
            selected = y.names[1])

selectInput(inputId="X", 
            label = strong("Predictor Variable: X"), 
            choices = x.names[], 
            selected = x.names[2])
      ###
sliderInput("newX", strong("New Value for Prediction:"), 2.5, min = 0, max = 700, step = 0.5)

```


\

<hr>

```{r}
  workDat = function(){
      if (input$Team == "ARI") {
          workingData = batters[which(batters$Team == "ARI"),]
      } else if (input$Team == "ATL") {
          workingData = batters[which(batters$Team == "ATL"),]
      } else if (input$Team == "BAL") {
          workingData = batters[which(batters$Team == "BAL"),]
      } else if (input$Team == "BOS") {
          workingData = batters[which(batters$Team == "BOS"),]
      } else if (input$Team == "CHC") {
          workingData = batters[which(batters$Team == "CHC"),]
      } else if (input$Team == "CIN") {
          workingData = batters[which(batters$Team == "CIN"),]
      } else if (input$Team == "CLE") {
          workingData = batters[which(batters$Team == "CLE"),]
      } else if (input$Team == "COL") {
          workingData = batters[which(batters$Team == "COL"),]
      } else if (input$Team == "CWS") {
          workingData = batters[which(batters$Team == "CWS"),]
      } else if (input$Team == "DET") {
          workingData = batters[which(batters$Team == "DET"),]
      } else if (input$Team == "HOU") {
          workingData = batters[which(batters$Team == "HOU"),]
      } else if (input$Team == "KC") {
          workingData = batters[which(batters$Team == "KC"),]
      } else if (input$Team == "LAA") {
          workingData = batters[which(batters$Team == "LAA"),]
      } else if (input$Team == "LAD") {
          workingData = batters[which(batters$Team == "LAD"),]
      } else if (input$Team == "MIA") {
          workingData = batters[which(batters$Team == "MIA"),]
      } else if (input$Team == "MIL") {
          workingData = batters[which(batters$Team == "MIL"),]
      } else if (input$Team == "MIN") {
          workingData = batters[which(batters$Team == "MIN"),]
      } else if (input$Team == "NYM") {
          workingData = batters[which(batters$Team == "NYM"),]
      } else if (input$Team == "NYY") {
          workingData = batters[which(batters$Team == "NYY"),]
      } else if (input$Team == "OAK") {
          workingData = batters[which(batters$Team == "OAK"),]
      } else if (input$Team == "PHI") {
          workingData = batters[which(batters$Team == "PHI"),]
      } else if (input$Team == "PIT") {
          workingData = batters[which(batters$Team == "PIT"),]
      } else if (input$Team == "SD") {
          workingData = batters[which(batters$Team == "SD"),]
      } else if (input$Team == "SEA") {
          workingData = batters[which(batters$Team == "SEA"),]
      } else if (input$Team == "SF") {
          workingData = batters[which(batters$Team == "SF"),]
      } else if (input$Team == "STL") {
          workingData = batters[which(batters$Team == "STL"),]
      } else if (input$Team == "TB") {
          workingData = batters[which(batters$Team == "TB"),]
      } else if (input$Team == "TEX") {
          workingData = batters[which(batters$Team == "TEX"),]
      } else if (input$Team == "TOR") {
          workingData = batters[which(batters$Team == "TOR"),]
      } else if (input$Team == "WAS") {
          workingData = batters[which(batters$Team == "WAS"),]
      } else {
          workingData = batters
      }
   workingData 
  }
```



Column {data-width=800 .tabset .tabset-fade}
-------------------------------------------------------------------
    
### **Interactive Scatterplot** 

```{r}
renderPlotly({
plot_ly(data = workDat(),
        x =  ~workDat()[[input$X]], 
        y =  ~workDat()[[input$Y]],
        color = ~workDat()$Team, 
        hovertemplate = paste('<i><b>',input$Y,'<b></i>: %{y}',
                             '<br><b>',input$X,'</b>:  %{x}',
                             '<br><b>',input$Team,'</b>'),
        alpha  = 0.9,
        size = ~workDat()[[input$Y]],
        type = "scatter",
        mode = "markers")   %>%
    layout(#title = paste( input$X, "vs", input$Y), 
           plot_bgcolor = "#e5ecf6",
           margin = list(l = 20, r = 20, b = 80, t = 10),
           xaxis = list(title = paste(input$X)), 
           yaxis = list(title = paste(input$Y)), 
           legend = list(title=list(text='<b> MLB Team </b>'),
                         orientation = "r",
                         xanchor = "left")
        )
  })
```

### **Boxplots of (X) Distribution**

```{r fig.align='center', fig.width=4, fig.height=3}
renderPlot({
 ggplot(data=workDat(), aes(x=workDat()$Team, y=workDat()[[input$X]])) +
        geom_boxplot(aes(fill=workDat()$Team)) +  
        ylab(input$X) + 
        xlab("MLB Team") + 
        labs(fill = "Team") +
        stat_summary(fun=mean, geom="point", shape=5, size=4)  
})
```

### **Regression Plot** 

```{r}
renderPlot({
  regdata = workDat()
  if (input$Team == "All"){
    m0 = lm(batters[[input$Y]]~batters[[input$X]])
    plot(batters[[input$X]], batters[[input$Y]],
      main = "", type = "p", pch = 19, col = "blue", xlab = input$X, ylab = input$Y)
    abline(m0, lwd = 2, col = "red")
  } else {
    m1 = lm(workDat()[[input$Y]]~workDat()[[input$X]])
    plot(workDat()[[input$X]], workDat()[[input$Y]],
      main = "", type = "p", pch = 19, col = "blue", xlab = input$X, ylab = input$Y)
    abline(m1, lwd = 2, col = "red")    
  }
})
```

### **Residual Plots**

```{r}
renderPlot({
  regdata = workDat()
  if (input$Team == "All"){
    LM = lm(batters[[input$Y]]~batters[[input$X]])
  } else {
    LM = lm(workDat()[[input$Y]]~workDat()[[input$X]])
  }
  par(mfrow = c(2,2))
  plot(LM)
})
```

### **Prediction Plot**

```{r}
renderPlot({
    dataset = workDat()

    ###
    m3 = lm(dataset[,input$Y] ~ dataset[,input$X])

    pred.y = coef(m3)[1] + coef(m3)[2]*input$newX
         #####
    plot(dataset[,input$X], dataset[,input$Y], 
         xlab = input$X, ylab = input$Y, main = "")
        ## adding a regression line to the plot
     abline(m3, col = "red", lwd = 1, lty=2)
    points(input$newX, pred.y, pch = 19, col = "red", cex = 2)
  })
```

### **Regression Coefficients**

```{r}
renderTable({
    dataset = workDat()
    m0 = lm(dataset[,input$Y] ~ dataset[,input$X])
    regcoef = data.frame(coef(summary(m0)))
    ##
    regcoef$Pvalue = regcoef[,names(regcoef)[4]]
    ###
    regcoef$Variable = c("Intercept", input$X)
    regcoef[,c(6, 1:3, 5)]
  })
```
