---
title: "Probability Distribution RShiny App"
author: "Andrew Heneghan"
date: "4/17/2024"
output: 
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: no
    code_folding: hide
    code_download: yes
    smooth_scroll: yes
    number_sections: yes
    theme: readable
---
      
<style type="text/css">
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0;
}
h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
}
h4.author { 
  /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}
h3 { 
    /* Header 3 - and the author and data headers use this too  */
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    font-face: bold;
    text-align: left;
}
h4 { 
    /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}
   /* center maps using chunk option: fig.align='center' */
.html-widget {
    margin: auto;
}
</style>

```{r setup, include=FALSE}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("shiny")) {
   install.packages("shiny")
   library(shiny)
}
if (!require("rsconnect")) {
   install.packages("rsconnect")
   library(rsconnect)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      fig.align='center', 
                      fig.pos = 'ht')
```

```{r}
library(shiny)
ui <- fluidPage(
  withMathJax(),
  tags$div(HTML("<script type='text/x-mathjax-config' >
            MathJax.Hub.Config({
            tex2jax: {inlineMath: [['$','$'], ['\\(','\\)']]}
            });
            </script >
            ")),
  ### 
  br(), br(), 
  wellPanel(style = "background-color:lightskyblue;",  
            # App title ----
            titlePanel(h2("Simulating Binomial Distribution", align = "center",
                          style = "color:midnightblue; font-family:verdana; 
                      font-variant: small-caps; font-weight: bold;"))),
  ###
  sidebarLayout(    # siderbarLayout:       
    sidebarPanel( style = "background-color:lightskyblue;", 
                  tags$div( h4("User Input Panel", align = "center",
                               style = "color:darkred; font-family:verdana; 
                      font-variant: small-caps; font-weight: bold;")),
                  #
                  br(),
                  ##
                  # slider input: number of experiments
                  sliderInput(inputId = "x",
                              label = "Number of Experiments",
                              min = 1,
                              max = 1000,
                              value = 500),
                  # slider input: number of trials per experiment
                  sliderInput(inputId = "n",
                              label = "Number of Trials",
                              min = 1,
                              max = 100,
                              value = 50),
                  # slider input: probability of success
                  sliderInput(inputId = "p",
                              label = "Probability of Success",
                              min = 0,
                              max = 1,
                              value = 0.5)),
    mainPanel( style = "background-color:khaki1; text-align: center;", 
               plotOutput(outputId = "distPlot")
    )
  ))
###
server <- function(input, output) {
  output$distPlot <- renderPlot({
    ##
    binom.score <- rbinom(input$x, input$n, input$p)
    ##
    par(bg = "khaki1")
    hist(binom.score, probability = TRUE, col = "deeppink", border = "darkgreen",
         xlab = "Histogram of Generated Binomial Data",
         main = "Simulated Binomial Distribution")
    d <- density(binom.score) # returns the density data
    lines(d)
  }, height = 400, width = 800)
}
###
shinyApp(ui = ui, server = server)
```



