---
title: "Penguin Data Analysis Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: cerulean
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

<style type="text/css">
/* Get a fancy font from Google Fonts */
@import url('https://fonts.googleapis.com/css2?family=Yusei+Magic&display=swap');

body {
  background-color: lightgray;
  color: lightsalmon; /* text color */
  text-align: left;
}

.navbar {
	position: fixed;
	display: flex;
	justify-content: center;
	width: 100%;
}

/* Make text visible on inputs */
.shiny-input-container {
  color: darkblue;
}

.navbg {
  background-color: blue;
  color: white;
}

</style>                    

\




```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(ggplot2)
library(tidyverse)
```

```{r data}
penguin1 = na.omit(read.csv(file="https://raw.githubusercontent.com/azheneghan/aheneghan/main/datasets/w03-penguins.csv"))[,-1]
penguin1$BMI = (penguin1$body_mass_g/4000)
x.names = names(penguin1)[c(3,4,5,9)]
y.names = names(penguin1)[c(3,4,5,9)]
location = unique(penguin1$island)
```


\


Column {.sidebar data-width=200}
-------------------------------------------------------------------

\

<hr>


```{r}
radioButtons(inputId = "island",
            label = strong("Location"), 
            choices = c("All", as.vector(location)), 
            selected = "All")
```

<hr>

```{r}
selectInput(inputId="Y", 
            label = strong("Response Variable: Y"), 
            choices = y.names[], 
            selected = y.names[1])

selectInput(inputId="X", 
            label = strong("Predictor Variable: X"), 
            choices = x.names[], 
            selected = x.names[2])
      ###
sliderInput(inputId = "newX", label = "New Value for Prediction:", value = 117, min = 0, max = 235, step = 0.1)

```


\

<hr>

```{r}
  workDat = function(){
    if (input$island == "Biscoe") {
      workingData = penguin1[which(penguin1$island == "Biscoe"),]
    } else if (input$island == "Dream") {
      workingData = penguin1[which(penguin1$island == "Dream"),]
    } else if (input$island == "Torgersen") {
      workingData = penguin1[which(penguin1$island == "Torgersen"),]
    } else {
      workingData = penguin1
    }
    workingData 
  }
```



Column {data-width=450 .tabset .tabset-fade}
-------------------------------------------------------------------

### **Scatterplot** 
```{r}
renderPlotly({
# hist(workDat()[[input$X]]) 
plot_ly(data = workDat(),
        x =  ~workDat()[[input$X]], 
        y =  ~workDat()[[input$Y]], 
        color = ~workDat()$island,
        colors = c("purple4", "darkorange", "indianred2"),
        hovertemplate = paste('<i><b>',input$Y,'<b></i>: %{y}',
                             '<br><b>',input$X,'</b>:  %{x}',
                             '<br><b>',input$island,'</b>'),
        alpha  = 0.9,
        size = ~workDat()[[input$Y]],
        type = "scatter",
        mode = "markers",
        width=500, height=450)   %>%
    layout(plot_bgcolor = "#e5ecf6",
           margin = list(l = 20, r = 20, b = 110, t = 20),
           xaxis = list(title = paste(input$X)), 
           yaxis = list(title = paste(input$Y)), 
           legend = list(title=list(text='<b> Location </b>'),
                         orientation = "r",
                         xanchor = "center")
        )
  })
```


### **Regression** 

```{r}
renderPlot({
  regdata = workDat()
  if (input$island == "All"){
    m0 = lm(penguin1[[input$Y]]~penguin1[[input$X]])
    plot(penguin1[[input$X]], penguin1[[input$Y]],
         main = paste("Relationship between", input$Y, "and", input$X),
         type = "p", pch = 19, col = "blue", xlab = input$X, ylab = input$Y)
    abline(m0, lwd = 2, col = "red")
  } else {
    m1 = lm(workDat()[[input$Y]]~workDat()[[input$X]])
    plot(workDat()[[input$X]], workDat()[[input$Y]],
         main = paste("Relationship between", input$Y, "and", input$X),
         type = "p", pch = 19, col = "blue", xlab = input$X, ylab = input$Y)
    abline(m1, lwd = 2, col = "red")    
  }
})
```

### **Residual Plots**

```{r}
renderPlot({
  regdata = workDat()
  if (input$island == "All"){
    LM = lm(penguin1[[input$Y]]~penguin1[[input$X]])
  } else {
    LM = lm(workDat()[[input$Y]]~workDat()[[input$X]])
  }
  par(mfrow = c(2,2))
  plot(LM)
})
```

### **Prediction**

```{r}
renderPlot({
    dataset = workDat()
    ###
    m3 = lm(dataset[,input$Y] ~ dataset[,input$X])
    ###
    pred.y = coef(m3)[1] + coef(m3)[2]*input$newX
         #####
    plot(dataset[,input$X], dataset[,input$Y], 
         xlab = input$X,
         ylab = input$Y,
         main = paste("Relationship between", input$Y, "and", input$X)
       )
    abline(m3, col = "red", lwd = 1, lty=2)
    points(input$newX, pred.y, pch = 19, col = "red", cex = 2)
  })
```



Column {data-width=400}
-------------------------------------------------------------------


### **Boxplots of Explanatory Variable by Location**
```{r}
renderPlot({
 #plot(workDat()[[input$X]], workDat()[[input$Y]]) 
 b <- ggplot(data=workDat(), aes(x=workDat()$island, y=workDat()[[input$X]])) +
        geom_boxplot(aes(fill=workDat()$island)) +  
        ylab(input$X) + 
        xlab("Location") + 
        #ggtitle(paste("Boxplot of the response variable: ", input$X)) +
        labs(fill = "Location") +
        stat_summary(fun.y=mean, geom="point", shape=5, size=4)
  b+scale_fill_manual(values=c("darkorange", "purple4", "indianred2"))
})
```


### **Ridge Plot for Distribution of Response by Location**
```{r}
renderPlotly({
  if(input$island == "All"){
    flipper.len.biscoe <- penguin1[which(penguin1$island == "Biscoe"),]
    biscoe <- density(flipper.len.biscoe$flipper_length_mm)
    flipper.len.dream <- penguin1[which(penguin1$island == "Dream"),]
    dream <- density(flipper.len.dream$flipper_length_mm)
    flipper.len.torgersen <- penguin1[which(penguin1$island == "Torgersen"),]
    torgersen <- density(flipper.len.torgersen$flipper_length_mm)
    # plot density curves
    plot_ly(x = ~torgersen$x, y = ~torgersen$y, 
               type = 'scatter',
               mode = 'lines', 
               name = 'Torgersen',
               line = list(color = 'indianred'),
               fill = 'toself')  %>% 
       add_trace(x = ~biscoe$x, y = ~biscoe$y, name = 'Biscoe', line = list(color = 'purple') , fill = 'toself')  %>%
       add_trace(x = ~dream$x, y = ~dream$y, name = 'Dream', line = list(color = 'darkorange'), fill = 'toself')  %>%
       layout(xaxis = list(title = paste(input$Y)), yaxis = list(title = 'Density'))
    } else {
      den <- density(workDat()[[input$Y]])
      plot_ly(x = ~den$x,
              y = ~den$y,
              color = ~input$island,
              colors = c("indianred2", "darkorange", "purple4"),
              type = 'scatter',
              mode = 'lines',
              name = input$island,
              fill = 'toself')    %>%
      layout(xaxis = list(title = paste(input$Y)), yaxis = list(title = 'Density'))
    }
})
```